---
title: "Adding pkgdown to your Package"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{adding-pkgdown}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Getting started

1. Open your package's `.Rproj` file

### Option 1

2. Run `usethis::use_pkgdown()` once to configure. This will generate additional files and sub-folders (e.g. `docs`).
3. Run `pkgdown::build_site()` to preview site and before any push to GitHub. There will be a lot of ancillary files, but check them in. They're all necessary for GitHub to build the website.

### Option 2

With this option, you don't need to activate Github Pages (in the next section) because it'll already be done.

```{r eval=FALSE}
usethis::use_pkgdown_github_pages()
pkgdown::build_site_github_pages()
```

Any .Rmd in the `vignette` sub-folder will be automatically converted to an article.

## Activate GitHub Pages:

1. Navigate in GitHub to `Settings` > `Pages`
2. Set the GitHub Pages Source to `main` branch & the sub-folder `docs`

It may take anywhere from 5 mins to 20 mins to see the website.

## Additional Information

https://pkgdown.r-lib.org/articles/pkgdown.html
