---
title: "Introduction"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{Introduction}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  collapse = TRUE,
  comment = "#>"
)
```

```{r setup}
library(mycensus)
```

```{r include = FALSE}
Sys.getenv("CENSUS_API_KEY")
```

`get_decennial_recs()` is an all purpose function to generate decennial estimates (2010 or beyond) for Census Tract, Counties, MSAs, and Census Places in Washington State. 

- Counties will default to King, Kitsap, Pierce, and Snohomish Counties. Otherwise provide a single county or a vector of counties
- Provide a single table code or a vector of table codes
- Provide a single FIPS code or a vector of FIPS codes when querying for MSAs or Census Places

Arguments can look like this:
```{r eval = FALSE}
Sys.getenv("CENSUS_API_KEY")
tbl_names <- paste0('PCT020', LETTERS[1:6])
get_decennial_recs(geography = 'county', table_codes = tbl_names, year = 2010)

get_decennial_recs(geography = 'tract', table_codes = tbl_names, year = 2010)

get_decennial_recs(geography = 'msa',
                  table_codes = c("H001", "P001"),
                  year = 2010,
                  fips = c('42660', "28420"))

```

If we queried for Census Places such as Seattle City and Burien City, results would look like this:
```{r eval = FALSE}
get_decennial_recs(geography = 'place',
                   table_codes = 'PCT013',
                   year = 2010,
                   fips = c("5363000", "5308850"))
```

```{r echo = FALSE, message=FALSE}
df <- get_decennial_recs(geography = 'place',
                         table_codes = 'PCT013',
                         year = 2010,
                         fips = c("5363000", "5308850"))
knitr::kable(head(df, 8))
```






