---
title: "Creating an R Package"
output: 
  html_document:
    theme: readable
    toc: true
    toc_depth: 3
    toc_float: true
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

My toy example and what I will reference throughout this document is here: https://github.com/cyyllam/mycensus

This document is a collection of notes and instructions gathered from the following sources: 

  - [DataCamp's tutorial on package creation](https://learn.datacamp.com/courses/developing-r-packages)
  - https://tinyheero.github.io/jekyll/update/2015/07/26/making-your-first-R-package.html

## Pre-work

Creating a package will make use of the following packages:  

  - `devtools`: contains tools to document, check, test, and install your package.
  - `roxygen2`: documentation headers above your functions that will help create R documentation when the user asks for help. 
  - `usethis`: automates workflow e.g. providing functions to insert information into a file or telling testthat to set up a testing framework for unit testing.
  - `testthat`: contains tools e.g. `expect_*()` to conduct unit tests.

Depending on what you're doing, [Rtools](https://cran.r-project.org/bin/windows/Rtools/) may be required. 

## Framework Set-up

Create package using RStudio's template. This creates a directory with an .Rproj file so your working directory is already set-up.

- Click on `New R Project` icon > `New Directory` > `New Package`

The name of your package should only contain letters, numbers, or dots.

### Description File

The DESCRIPTION file contains author and maintainer details, package dependencies, and license details.

To add a package dependency, you can use `usethis::use_package()`. It will automatically insert whatever package you provide.

`cre` = maintainer (who should get an email if something goes wrong)

### NAMESPACE File

The NAMESPACE file that exists when you create your package directory is a minimalist file. Delete it! 

This file will be re-created and updated when you run the `devtools::document()` after completing documentation on your functions.

The NAMESPACE file is something that you do not edit by hand. 

## Adding Functions & roxygen

Add functions in one or across multiple R files in the `R` directory.

Place cursor in the body of your function. Use `Ctrl` + `Alt` + `Shift` + `R` to add Roxygen2 commenting structure and edit accordingly.

Examples:

  - See an example from my toy package: [mycensus](https://github.com/cyyllam/mycensus/blob/792c857c91871eaf0d95f513fc0451ef1483b2cf/R/get-decennial-recs.R#L141)
  - https://github.com/kbroman/pkg_primer/blob/gh-pages/example/stage3/R/brocolors.R

You can run `devtools::document()` to convert scripts with roxygen headers to .Rd files. These are the R Documentation Help files when use `?` or `help()` to check function details.

`@export` is used for functions you want to be made available to the user. It will export the function to the NAMESPACE and when the user starts typing `mycensus::` only exported functions will appear in the intellesense. 

To document the package itself, you can store the following at the top of a script with roxygen headers. It can also be in a separate file.

```{r eval = FALSE}
#' mycensus: A package for easy compilation of census estimates for the Puget Sound Region.
#'
#' This package provides functions to easily collect census estimates across geographies within the
#' Central Puget Sound region.
#'
#' @author Christy Lam \email{clam@psrc.org}
#' @docType package
#' @name mycensus
"_PACKAGE"

utils::globalVariables("GEOID")
```

Using variable names in dplyr functions may trigger warnings or notes on undefined variables. Use the `globalVariables` at the top of your script(s) to pass the checking stage.

## R CMD Check

Run `devtools::check()` to check package for errors, warnings, and notes. Make sure to resolve all errors at the very least before moving on.

`check()` will automatically update .Rd files.

## Distribution

After error checking and completing the DESCRIPTION file, add/commit/push to Github.

Install your package on your computer:
```{r eval = FALSE}
devtools::install()
library(mycensus)
```

To distribute your package via Github have others run this command:
```{r eval = FALSE}
devtools::install_github("cyyllam/mycensus")
```

Another way of distributing your package is to create a source file (a `.tar.gz` file). Use `devtools::build()` and it will generate a file outside of your package directory (I'm not sure where to store the file though...).

## Unit Tests

Set-up test framework using `usethis::use_testthat()`.

Save your tests as .R files in the `tests/testthat/` directory with the prefix `test-`.

Use various `expect_*()` functions to conduct tests.

To group tests (for organization), use the function `testthat::test_that()` like so:
```{r eval = FALSE}
testthat::test_that('get_deccenial_recs returns expected results', {
  df <- mycensus::get_decennial_recs(geography = 'county',
                                     table_codes = 'PCT013',
                                     year = 2010)

  # default counties are formatted with 'County, Washington'
  testthat::expect_identical(unique(df$NAME), c('King County, Washington', 'Kitsap County, Washington', 'Pierce County, Washington', 'Snohomish County, Washington'))

  # default columns from tidycensus are the follow six variables
  testthat::expect_identical(colnames(df), c('GEOID', 'NAME', 'variable', 'value', 'label', 'concept'))

})

```

To run all tests: `devtools:test()`

## Continuous Integration

Consider using Travis CI (continuous integration) or `usethis::use_github_actions()` (which I haven't done yet). Every time code is committed/pushed to Github, the R CMD Check and unit tests will automatically run.

For more reading:
https://juliasilge.com/blog/beginners-guide-to-travis/ (this was written back in 2016) and `use_travis()` is now soft-deprecated. Documentation recommends `usethis::use_github_actions()` for CI.

## Explore!

View the vignette for [mycensus](Introduction.html) using `use_vignette()`, but a better way to document packages might be through the [`pkgdown`](https://pkgdown.r-lib.org/) package.

```{r eval = FALSE}
library(mycensus)

# check out the documentation!
?mycensus
?get_decennial_recs
```

